Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-05-07 doc/GMP2015/revision_2/s02_preliminaries.tex
Major changes to paper (revision 2)

Basics definitions and theory for this topic can be found in~\ref{sec:appendix:defs}.

@@

@@

In this paper, a \dfn{qusecs} is any independent geometric constraint system of one of 3 types: bar-joint (defined formally in \ref{sec:appendix:defs}), \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}), and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).

@@

\ClearMyMinHeight
\SetMyMinHeight{.4}{img/epsfromtikz/demo_graph}
\SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_comdrp}
\SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_candrp}

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.4\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
    \caption{}\label{fig:demo_graph:graph}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
    \caption{}\label{fig:demo_graph:comdrp}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
    \caption{}\label{fig:demo_graph:candrp}
  \end{subfigure}%
  %
  \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$. The children of triangles are omitted.}
  \label{fig:demo_graph}
\end{figure*}%

\begin{figure*}\centering%
  %
  \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
  \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality. Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
  \label{fig:demo_graph:clustmindrp}
\end{figure*}%

>>>>

% Basics definitions and theory for this topic can be found in~\ref{sec:appendix:defs}.

@@

\subsection{Basic Definitions and Theory}
\label{sec:prelim:defs}
\label{sec:appendix:defs}

In this paper, a \dfn{geometric constraint system} is a multivariate polynomial (usually bilinear or quadratic) system $G(x,\delta)=0$, representing constraints with parameters $\delta$ between geometric primitives  in $\mathbb{R}^2$ represented collectively as $x\in \mathbb{R}^n$.
%
When the type of constraint (system) is fixed, the system is simply represented as $(G,\delta)$, where $G$ is the underlying constraint (hyper)graph $G = (V,E)$ with the vertices $V$ representing the geometric primitives in $\mathbb{R}^2$ and (hyper)edges $E$ representing the constraints, each with an associated parameter $\delta$.
%
For example, a \dfn{bar-joint system or linkage} $(G,\delta)$, is a graph $G=(V,E)$ with fixed length bars as edges,
% \footnote{Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of values rather than exact values.}
i.e. $\delta: E \rightarrow \mathbb{R}$; this represents the distance constraint system $\| x_u -x_v \|_2 = \delta_{u,v}$ for  $(u,v) \in E$, where $x_u \in \mathbb{R}^2$ represents the coordinates of $u\in V$.

In this paper, we will consider 3 types of constraint systems: bar-joint as well as \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}) and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}). In all of these cases, a Cartesian \dfn{realization} or \dfn{solution} $G(p)$ of $(G,\delta)$ is an assignment of coordinates or Euclidean transformations (poses), $p: V \rightarrow \mathbb{R}^2$ or $\mathbb{R}^3$, to the vertices of $G$ satisfying the constraints with parameters $\delta$, modulo orientation preserving isometries (Euclidean rigid body motions for distance constraint).

Although the realization space itself depends on the constraint parameters $\delta$, many relevant \dfn{generic} properties of the constraint system $G(x,\delta)$ are properties of the constraint (hyper)graph $G$ and do not depend on $\delta$ (or they hold for all but a measure zero set of $\delta$ values). Many of these are properties of the Jacobian $\Delta_x G(x,\delta)$, often called the appropriate \dfn{rigidity matrix of $G$} (a matrix of indeterminates). For example, the \dfn{bar-joint rigidity matrix of the graph $G = (V,E)$} is a matrix of indeterminates representing the Jacobian of the distance map $\| x_u -x_v \|_2$  for $(u,v) \in E$. The matrix has $2$ columns per vertex in $V$ and one row per edge in $E$, where the row corresponding to edge $(u,v)$ contains the 2 coordinate indeterminates for $x_u -x_v$ (resp. $x_v-x_u$) in the 2 columns for $u$ (resp.\ $v$), that is 4 non-zero entries per row.

% XXXXXXXXX
% something has to be said along with asimow roth
% (infinitesimal rigidity implies rigidity) When the rigidity matrix has
% appropriate rank, the realizations or solutions of the corresponding
% constraint system are generically isolated and zero-dimensional if the
% constraint system has equalities and exact values for the parameters
% $\delta$. If the constraint system has inequalities, i.e, if the
% $\delta$ lie in an interval, this claim is approximate: the solutions
% are isolated small, full-dimensional  connected components.

One important property of generic constraint system or (hyper)graph\footnote{We refer to these as properties of the constraint system or as properties of the underlying (hyper)graph interchangeably.}
%that depends on the appropriate rigidity matrix of indeterminates
is \dfn{rigidity},
% (\note footnote now...),
i.e., the realizations or solutions of the corresponding constraint system being generically isolated and zero-dimensional.%(if the constraint system has equalities and exact values for the parameters $\delta$).
The result by Asimow and Roth \cite{asimow1978rigidity}  shows a constraint (hyper)graph is rigid if it is generically \dfn{infinitesimally rigid}, i.e.\ the number of independent rows of its appropriate rigidity matrix is at least the number of columns less the number of rigid body motions, which is 3 for bar-joint systems.
%This condition is called generic {\em infinitesimal rigidity}.

% \sidenote{If the constraint system has inequalities, i.e, if the $\delta$ lie in an interval, the definition of rigidity is approximate: the solutions are isolated small, full-dimensional  connected components.}
% generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system

Geometric constraint systems can also have inequalities in addition to equations, where the parameters in $\delta$ are small intervals of rather than exact values. In this case, the definition of rigidity is approximate; the solutions are isolated, small, full-dimensional connected components.

Other generic constraint system or (hyper)graph properties are the following:
% (\note we refer to these as properties of the constraint
% system or as properties of the underlying (hyper)graph
% interchangeably).
A constraint (hyper)graph $G$ is \dfn{independent} if its appropriate rigidity matrix of indeterminates has independent rows (i.e, the determinant of some square submatrix is not identically zero).
%It is {\em rigid} if the number of independent rows of the rigidity matrix is
%at least the number of columns less the number of rigid body motions,
%which is 3 for distance constraint systems.
It is \dfn{isostatic (minimally rigid, wellconstrained)} if it is both rigid and independent.
%if the number of generically independent rows or the rank of the appropriate rigidity matrix is maximal.
%For example, the maximal rank of a bar-joint rigidity matrix is $2|V| - 3$,
%where $3$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^2$.
%The graph $G$ is {\em rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is wellconstrained.
It is \dfn{flexible} if it is not rigid, \dfn{underconstrained} if it is independent and not rigid, or \dfn{overconstrained} if it is not independent.

Defining the combinatorial independence of a subset of edges $E'\subseteq E$ to be the generic independence of corresponding rows in the appropriate rigidity matrix of indeterminates, we obtain the \dfn{rigidity matroid} of a constraint (hyper)graph $G = (V,E)$.
%The 2-dimensional {\em rigidity matroid} of a constraint (hyper)graph $G = (V,E)$ is a linear matroid  on $E$,
%where a subset of edges $E' \subseteq E$ is {\em independent} in the matroid,
%if the set of corresponding rows in the appropriate rigidity matrix of indeterminates are linearly independent.
There are various results on combinatorial characterization of independence and rigidity  and rigidity matroids for different types of graphs. For bar-joint rigidity matroid, the famous Laman's theorem \cite{laman1970graphs} states that the underlying graph is isostatic if and only if $|E| = 2|V|-3$ and $|E'| \le 2|V'|-3$ for every induced subgraph with at least 2 vertices. The result by Lovasz and Yemini \cite{lovasz1982generic}  shows that all 6-vertex-connected graphs are rigid in the plane. For bar-body rigidity matroid, Tay \cite{tay1976rigidity} proved that the underlying multigraph is isostatic if and only if it can be decomposed as $3$ edge disjoint spanning trees. White and Whiteley \cite{white1987algebraic} gave the same characterization using a different technique to study the algebraic-geometric conditions of genericity, called pure condition. Lee, Streinu and Theran \cite{lee2007graded} defined the \dfn{$(k,l)$-sparsity matroid}, where a hypergraph $G$ is called \dfn{$(k,l)$-sparse} if $|E'| \le k|V'| - l$ for any induced subgraph $(V',E')$ with at least 2 vertices, and \dfn{$(k,l)$-tight} if is $(k,l)$-sparse and $|E| = k|V| - l$. In general, given a $d$-uniform hypergraph, a $(k,l)$-sparsity condition is matroidal as long as $l \le dk-1$.

% In this paper, a \dfn{qusecs} is any \dfn{independent} geometric constraint system of one of the 3 types mentioned above.

% % \medskip\noindent
% % \note In the remainder of this section and sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs. Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications.

% \sidenote{In the remainder of this section and Sections~\ref{sec:DRP} and \ref{sec:recomb} we only consider bar-joint qusecs and graphs. Relevant formal analogies for the other 2 types of qusecs and (hyper)graphs are given in the subsequent 2 sections on Applications (\ref{sec:bodypin} and \ref{sec:pinnedline}).}

%XXXNow comes the definition of stress vector and flex vector.
%NOTE: something has to be said about what flexes and stresses mean

%TODO
Given a bar-joint graph, a vector (of indeterminates) in the right null space of its rigidity matrix is called a \dfn{flex vector}. It has 2 entries per vertex and represents the internal motions of the system (i.e.\ modulo rigid body motions). For rigid graphs, all flex vectors are identically zero.
%
A vector (of indeterminates) in the left null space of the rigidity matrix is called a \dfn{stress vector}. It has one entry per edge and represents a \dfn{self stress} of the system. For independent graphs, all stress vectors are identically zero.
%
An \dfn{external stress} $t$ is a $1\times 2|V|$ vector  (of indeterminates) which specifies a 2D vector acting at each vertex such that the stress balance equation $sR = t$ holds, where $s$ is a vector of internal stresses, one for each bar, and $R$ is the rigidity matrix.

@@

In this paper, a \dfn{qusecs} is any independent geometric constraint system of one of 3 types: bar-joint (defined formally in \ref{sec:prelim:defs}), \dfn{body-hyperpin} (defined formally in Section \ref{sec:bodypin}), and \dfn{pinned line-incidence} (defined formally in Section \ref{sec:pinnedline}).

@@

% \ClearMyMinHeight
% \SetMyMinHeight{.4}{img/epsfromtikz/demo_graph}
% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_comdrp}
% \SetMyMinHeight{.3}{img/epsfromtikz/demo_graph_candrp}

% \begin{figure*}\centering%
%   %
%   \begin{subfigure}{0.4\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph}
%     \caption{}\label{fig:demo_graph:graph}
%   \end{subfigure}%
%   %
%   \hfill
%   \begin{subfigure}{0.3\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_comdrp}
%     \caption{}\label{fig:demo_graph:comdrp}
%   \end{subfigure}%
%   %
%   \hfill
%   \begin{subfigure}{0.3\linewidth}\centering
%     \includegraphics[height=\myMinHeight]{img/epsfromtikz/demo_graph_candrp}
%     \caption{}\label{fig:demo_graph:candrp}
%   \end{subfigure}%
%   %
%   \caption{(\ref{fig:demo_graph:graph}) A simple graph, $G_{\Delta}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{\Delta}$, i.e.\ $ComDRP(G_{\Delta})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{\Delta}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{\Delta})$. The children of triangles are omitted.}
%   \label{fig:demo_graph}
% \end{figure*}%

% \begin{figure*}\centering%
%   %
%   \includegraphics[width=0.4\linewidth]{img/epsfromtikz/demo_graph_clustmindrp}
%   \caption{A DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which gives cluster minimality. Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the optimal DR-plans shown in Figure~\ref{fig:demo_graph:candrp} and \ref{fig:demo_graph:candrpseq} have fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
%   \label{fig:demo_graph:clustmindrp}
% \end{figure*}%

\ClearMyMinHeight
\SetMyMinHeight{.4}{img/svg/3xc2c3}
\SetMyMinHeight{.3}{img/svg/3xc2c3_comdrp}
\SetMyMinHeight{.3}{img/svg/3xc2c3_candrp}

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.4\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3}
    \caption{}\label{fig:demo_graph:graph}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3_comdrp}
    \caption{}\label{fig:demo_graph:comdrp}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/3xc2c3_candrp}
    \caption{}\label{fig:demo_graph:candrp}
  \end{subfigure}%
  %
  \caption{(\ref{fig:demo_graph:graph}) A graph, $G_{demo}$, used to illustrate concepts throughout this and the next section. (\ref{fig:demo_graph:comdrp}) The complete DR-plan of $G_{demo}$, i.e.\ $ComDRP(G_{demo})$. Dashed lines indicate that the children repeat the same pattern as the others shown on this level. The children of triangles (3 edges) are omitted. (\ref{fig:demo_graph:candrp}) The canonical DR-plan of $G_{demo}$, which is optimal (see Section~\ref{sec:DRP}), i.e.\ $OptDRP(G_{demo})$. The children of triangles are omitted.}
  \label{fig:demo_graph}
\end{figure*}%

\ClearMyMinHeight
\SetMyMinHeight{.4}{img/svg/triangle}
\SetMyMinHeight{.3}{img/svg/triangle_candrp}
\SetMyMinHeight{.3}{img/svg/triangle_clustmindrp}

\begin{figure*}\centering%
  %
  \begin{subfigure}{0.4\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/triangle}
    \caption{}\label{fig:demo_graph_tri:graph}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/triangle_candrp}
    \caption{}\label{fig:demo_graph_tri:candrp}
  \end{subfigure}%
  %
  \hfill
  \begin{subfigure}{0.3\linewidth}\centering
    \includegraphics[height=\myMinHeight]{img/svg/triangle_clustmindrp}
    \caption{}\label{fig:demo_graph_tri:clustmindrp}
  \end{subfigure}%
  %
  \caption{(\ref{fig:demo_graph_tri:graph}) A simple graph, $G_{tri}$, of 3 triangles intersecting trivially. (\ref{fig:demo_graph_tri:candrp}) The canonical DR-plan of $G_{tri}$. (\ref{fig:demo_graph_tri:clustmindrp}) A DR-plan of graph $G_{tri}$ which gives cluster minimality. Using the ideas from the paper~\cite{lomonosov2004graph}, the children of a triangle are the primitives of a point and an edge. Thus, the fan-in of this graph is 3, whereas the canonical/optimal DR-plan has fan-in of 2. With this counterexample, it is clear that cluster minimality is not sufficient for optimality.}
  \label{fig:demo_graph:clustmindrp}
  \label{fig:demo_graph_tri}
\end{figure*}%

### 2015-05-07 doc/GMP2015/revision_2/s03_drp.tex
Major changes to paper (revision 2)

In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic a complete DR-plan's uniqueness, while for independent (underconstrained or isostatic graphs), we show that the nonunique parts do not affect optimality.

\subsection{Previous Work}
\todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.}

@@

\subsection{Theory}
\sidenote{In this section, $G$ is always assumed to be isostatic.}

This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.

@@

This directly satisfies properties 2 and 3 of a canonical DR-plan (see definition \ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs. However, we still have to show property 1 (that this makes a DR-plan).
For rule (1), because we are beginning with a complete DR-plan, if we keep all the children it is still a DR-plan. For rule (2), we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we have found a larger rigid vertex-maximal proper subgraph.

@@

This proof of this theorem relied on the following crucial observation and lemma that are also used for other types of qusecs in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper.

@@

\begin{lemma*}\label{lemma:combined_lemma}
Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$
Then
\begin{enumerate}
    \item\label{lemma:wc_intersection_is_C}
    $C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.

@@

\begin{figure*}\centering%
  %
  \includegraphics[width=0.3\linewidth]{img/epsfromtikz/demo_graph_candrp_seq}
  \caption{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}). The children of the triangle are omitted. Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}. Also, note that the bottom-left node, the triangle $bce$, is the intersection of the 3 children of $G_{\Delta}$ in $ComDRP(G_{\Delta})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
  \label{fig:demo_graph:candrpseq}
\end{figure*}%

@@

\begin{proof}
The first step of the algorithm is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$. Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$. The output of this will be a list of component-candidates. Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$. If $D=G$, move the candidate (before adding $e$) to the true-component list and remove it from the candidate-component list. If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$ back to the list. Continue with the next candidate. When you exhaust the candidates, move to the next step.

Test the intersection of any two of the true-components. If the intersection is trivial, the entire list becomes the children. Recursively apply the algorithm to each child.

If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k\in S_N}{C_k}$. The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}, and we use the notation here. We construct a DR-plan whose node set is the same as that of a canonical/optimal DR-plan. Begin with the parent $C$. Its children will be $C\setminus R_1$ and $R_1$ plus the nodes and edges adjacent to $R_1$ that lie in $I$. Recursively apply the algorithm to the child containing $R_1$. The children of $C\setminus R_1$ will be $C\setminus (R_1\cup R_2)$ and $R_2$ plus the nodes and edges adjacent to $R_2$ that lie in $I$. And so on. At $N$ levels down, the 2 children will be $I$ and $R_N$ plus the nodes and edges adjacent to $R_N$ that lie in $I$. Apply the algorithm recursively to both children. See Figure~\ref{fig:demo_graph:candrpseq} for an example.
% Although the nodes containing $R_1,\ldots,R_N$ are underconstrained, simply take the children to be the roots of the forest

The reason for this construction is that the result is a tree with no duplicate nodes. The leaves of this tree will exactly be the edges of this graph. Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.

>>>>

In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the uniqueness of a complete DR-plan, and we show that the nonunique parts do not affect optimality for independent (underconstrained or isostatic) graphs. Furthermore, we give an efficient \candrpcomplexity\ algorithm to find the canonical DR-plan of any independent graph. The definition is as follows:

\subsection{Previous Work}
% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.}
We now briefly survey existing techniques for detecting rigidity and creating DR-plans of 2D constraint systems. The limitations of these techniques directly motivate the contributions in this section.

@@

\subsection{Theory}
In this section and in section~\ref{sec:recomb}, we deal exclusively with isostatic constraint systems. Any reference to some graph $G$ is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
% \sidenote{In this section, $G$ is always assumed to be isostatic.}

% This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.

Definition~\ref{def:canonical_drplan} gives the canonical DR-plan a surprisingly strong Church-Rosser property, which is made explicit in Theorem~\ref{theorem:main}, the main result of this section.

@@

This directly satisfies properties 2 and 3 of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs. However, we still have to show property 1 (that this makes a DR-plan).
For rule (1), because we are beginning with a complete DR-plan, if we keep all the children it is still a DR-plan. For rule (2), we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we have found a larger rigid vertex-maximal proper subgraph.

@@

This proof of this theorem relied on the following crucial observation and lemma. These will be used again in the application sections (\ref{sec:bodypin} and \ref{sec:pinnedline}) of the paper, with modifications to work for other types of qusecs.

@@

\begin{lemma*}\label{lemma:combined_lemma}
Let $C$ be a node of a canonical DR-plan, with distinct children $C_i$ and $C_j$.
Then
\begin{enumerate}
    \item\label{lemma:wc_intersection_is_C}
    $C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.

@@

% \begin{figure*}\centering%
%   %
%   \includegraphics[width=0.3\linewidth]{img/epsfromtikz/demo_graph_candrp_seq}
%   \caption{The sequential canonical DR-plan of $G_{\Delta}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}). The children of the triangle are omitted. Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}. Also, note that the bottom-left node, the triangle $bce$, is the intersection of the 3 children of $G_{\Delta}$ in $ComDRP(G_{\Delta})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
%   \label{fig:demo_graph:candrpseq}
% \end{figure*}%
\begin{figure*}\centering%
  %
  \includegraphics[width=0.3\linewidth]{img/svg/3xc2c3_candrp_seq}
  \caption{The sequential canonical DR-plan of $G_{demo}$ from Figure~\ref{fig:demo_graph:graph}, which is optimal (as explained in the proof of Theorem~\ref{theorem:algo_complexity}). The children of the triangle are omitted. Compare to to the typical canonical DR-plan shown in Figure~\ref{fig:demo_graph:candrp}. Also, note that the bottom-left node, the triangle, is the intersection of the 3 children of $G_{demo}$ in $ComDRP(G_{demo})$, shown in Figure~\ref{fig:demo_graph:comdrp}.}
  \label{fig:demo_graph:candrpseq}
\end{figure*}%

@@

\begin{proof}
The first step of the algorithm, which we call $CanDRP(G)$, is finding the isostatic vertex-maximal proper subgraphs of the input isostatic graph $G$. Do this by first dropping arbitrary edge $e$ from the edge set and running the pebble game algorithm \cite{Jacobs:1997:PG} on this subgraph, which is $O(|V|^2)$. The output of this will be a list of component-candidates. Then run the Frontier algorithm \cite{hoffman2001decompositionII} \cite{lomonosov2004graph} ($O(|V|)$) on each candidate along with edge $e$, which will find the minimal subgraph $D$ that contains the candidate and $e$. If $D=G$, move the candidate (before adding $e$) to the true-component list and remove it from the candidate-component list. If $D\subsetneq G$, remove all candidates that are subgraphs of $D$ from the candidate-component list and add $D$ back to the list. Continue with the next candidate. When you exhaust the candidates, move to the next step.

Test the intersection of any two of the true-components. If the intersection is trivial, the entire list become the children; recursively apply the algorithm to each child. If the intersection is non-trivial (necessarily isostatic), immediately compute the sequential decomposition down to $I=\bigcap_{k=1}^{N}{C_k}$. The root of this idea is discussed in detail in the proof of Theorem \ref{theorem:main}. We construct a DR-plan whose node set is the same as that of a canonical/optimal DR-plan. Begin with the parent $C$. Its children will be $S_1$ and $CanDRP(T_1)$, where $S_1=C\setminus R_1$ and $T_1$ is $R_1$ plus all incident edges (observe that these edges will add nodes from $I$ and that $T_1$ is underconstrained so it will be a forest and each root becomes a child of $C$). The children of $S_1$ will be $S_2$ and $CanDRP(T_2)$, where $S_2=C\setminus (R_1\cup R_2)$ and $T_2$ is $R_2$ plus all incident edges. And so on. At $N$ levels down, the children will be $S_N=I$ and $CanDRP(T_N)$, where $T_N$ is $R_N$ plus all incident edges. Apply the algorithm recursively to both children. See Figure~\ref{fig:demo_graph:candrpseq} for an example. \todo{should this more elaborate discussion be moved outside the proof?}

The previous discussion was a detailed explanation with an implied efficient method to solve for the case of non-trivial intersections, which avoids recomputation of the isostatic components of $C$. This stage of the algorithm could be stated more succinctly:
if the intersection is trivial, recursively apply $CanDRP$ to the entire list of true-components and set the resulting trees as the children. Else, the intersection is non-trivial, and the children of the node are $CanDRP(C_1)$ and $CanDRP(T)$, where $C_1$ is the first true-component of the parent $C$ (chosen arbitrarily, could be any child) and $T$ is the underconstrained graph formed from $C\setminus C_1$ plus all incident edges (also adding the associated nodes in $C_1$).

% Although the nodes containing $R_1,\ldots,R_N$ are underconstrained, simply take the children to be the roots of the forest

The reason for this construction is that it makes analysis of the algorithm convenient. The result of the algorithm is DR-plan that is a tree with no duplicate nodes. The leaves of this tree will be exactly the edges of this graph. Therefore, the number of nodes in the tree is $O(|E|)$ which, for isostatic input graphs, is $O(|V|)$.

### 2015-05-07 doc/GMP2015/revision_2/s04_recombination.tex
Major changes to paper (revision 2)

\subsection{Previous Work}
\todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.}

>>>>

\subsection{Previous Work}
% \todo{We now briefly survey existing techniques for studying 2D qusecs,  many of which are \dfn{bar-joint} systems (Examples 1 and 2 above, see Sections \ref{sec:prelim}, \ref{sec:DRP}, and \ref{sec:recomb}), \dfn{body-hyperpin} systems (Example 4 and 5, see Section \ref{sec:bodypin}) or \dfn{pinned-line incidence} systems (Example 3, see Section \ref{sec:pinnedline}). The limitations of these techniques directly motivate the contributions of this paper.}
We now briefly survey existing techniques for handling the complexity of recombination of 2D DR-plans. The limitations of these techniques directly motivate the contributions in this section.
